---
output: 
  html_document: 
    keep_md: yes
---
[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/pdp)](https://cran.r-project.org/package=pdp)
[![Build Status](https://travis-ci.org/bgreenwell/pdp.svg?branch=master)](https://travis-ci.org/bgreenwell/pdp)
[![Coverage Status](https://img.shields.io/codecov/c/github/bgreenwell/pdp.svg)](https://codecov.io/github/bgreenwell/pdp?branch=master)
[![Downloads](http://cranlogs.r-pkg.org/badges/pdp)](http://cranlogs.r-pkg.org/badges/pdp)
[![Total Downloads](http://cranlogs.r-pkg.org/badges/grand-total/pdp)](http://cranlogs.r-pkg.org/badges/grand-total/pdp)

pdp: An R Package for Constructing Partial Dependence Plots
================

The primary purpose of this package is to provide a general framework for constructing _partial dependence plots_ (PDPs) in R.


### Installation

The R package `pdp` is available from [CRAN](https://cran.r-project.org/package=pdp); the development version is hosted on [GitHub](https://github.com/bgreenwell/pdp). There are two ways to install:
```{r, eval=FALSE}
# Install latest release from CRAN (recommended)
install.packages("pdp")

# Install development version from GitHub repo
devtools::install_github("bgreenwell/pdp")
```


### Example usage

The examples below demonstrate various usages of the `pdp` package: regression, classification, and interfacing with the well-known `caret` package. To start, we need to install a few additional packages that will be required to run the examples.
```{r, eval=FALSE}
install.packages(c("caret", "ggmap", "kernlab", "randomForest"))
```


#### Regression example

In this example, we fit a random forest to the Boston housing data. Note, for any of the following examples, you can see a progress bar by simply specifying `.progress = "text"` in the call to `partial`.
```{r, message = FALSE, warning=FALSE}
# Load required packages
library(ggmap)  # for plotting data on maps
library(pdp)  # for constructing PDPs
library(randomForest)  # for random forest algorithm

# Load the Boston housing data
data (boston)  # included with the pdp package

# Fit a random forest to the boston housing data
set.seed(101)  # for reproducibility
boston.rf <- randomForest(cmedv ~ ., data = boston)

# Using randomForest's partialPlot function
partialPlot(boston.rf, pred.data = boston, x.var = "lstat")

# Package pdp's partial function returns a data frame by default
head(partial(boston.rf, pred.var = "lstat"))  # returns a data frame

# Set `plot = TRUE` to return a "trellis" object
partial(boston.rf, pred.var = "lstat", plot = TRUE, rug = TRUE)

# The partial function allows for multiple predictors
partial(boston.rf, pred.var = c("lstat", "rm"), grid.resolution = 10,
        plot = TRUE, chull = TRUE)

# The plotPartial function offers more flexible plotting
pd <- partial(boston.rf, pred.var = c("lstat", "rm"), grid.resolution = 10)

# Default plot
plotPartial(pd)

# Customized plot
plotPartial(pd, levelplot = FALSE, zlab = "cmedv", drape = TRUE,
            colorkey = FALSE, screen = list(z = -20, x = -60))

# Look at partial dependence of median home value on location
pd.loc <- partial(boston.rf, pred.var = c("lon", "lat"), chull = TRUE)

# Overlay predictions on a map of Boston
# ll <- c(range(boston$lon), range(boston$lat))
# map <- get_map(location = ll[c(1, 3, 2, 4)], zoom = 11, maptype = "toner-lite")
# ggmap(map) + 
#   geom_point(aes(x = lon, y = lat), data = boston, alpha = 0.2) +
#   geom_tile(aes(x = lon, y = lat, z = y, fill = y), 
#             data = pd.loc, alpha = 0.3) +
#   geom_contour(color = "white", alpha = 0.5) +
#   scale_fill_distiller(palette = "Spectral", name = "Median\nvalue") +
#   coord_fixed(ratio = 1) +
#   labs(x = "Longitude", y = "Latitude")
```


#### Classification example 

In this example, we fit another random forest to the Pima Indians diabetes data.
```{r, message=FALSE, warning=FALSE}
# Fit a random forest to the Pima Indians diabetes data
data (pima)  # load the boston housing data
set.seed(102)  # for reproducibility
pima.rf <- randomForest(diabetes ~ ., data = pima, na.action = na.omit)

# Partial dependence of glucose on diabetes test result (neg/pos)
partial(pima.rf, pred.var = c("glucose", "age"), plot = TRUE, chull = TRUE)
```


#### Interface with `caret`

Finally, we demonstrate the seemless interface between the `pdp` package and the `caret` package. For illustration we use `caret`'s `train` function to tune a support vector machine (SVM) with a radial basis function kernel to the Pima Indians diabtes data. We then cinstruct PDPs from the final model for the most "important" variables. When training a model using `caret`'s `train` function, you can view tuning progress by setting `verboseIter = TRUE` in the call to `trainControl`.
```{r, message=FALSE, warning=FALSE}
# Load required packages
library(caret)  # for model training/tuning

# Set up for 5-fold cross-validation
ctrl <- trainControl(method = "cv", number = 5)

# Tune a support vector machine (SVM) using a radial basis function kerel to
# the Pima Indians diabetes data
set.seed(103)  # for reproducibility
pima.svm <- train(diabetes ~ ., data = pima, method = "svmRadial",
                  prob.model = TRUE, na.action = na.omit, trControl = ctrl,
                  tuneLength = 10)

# Partial dependence of glucose on diabetes test result (neg/pos)
partial(pima.svm, pred.var = "glucose", plot = TRUE, rug = TRUE)
```
