---
title: "Bouston  Tests"
author: "Brandon M. Greenwell"
date: "July 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
# Load required packages
library(ggplot2)
library(grid)
library(pdp)

# Set knitr chunk options
knitr::opts_chunk$set(echo = TRUE)

# Function to run tests for a fitted model object using the iris data
irisTest <- function(object, ...) {
  
  # Training data
  trn <- subset(iris, select = -Species)
  
  # Partial dependence plot for Petal.Length on class 1 (i.e. setosa)
  pd1 <- partial(object, pred.var = "Petal.Length", which.class = 1, 
                 prob = TRUE, train = trn, ...)
  
  # Partial dependence plot for Petal.Length on class 2 (i.e. versicolor)
  pd2 <- partial(object, pred.var = "Petal.Length", which.class = 2, 
                 prob = TRUE, train = trn, ...)
  
  # Partial dependence plot for Petal.Length on class 3 (i.e. verginica)
  pd3 <- partial(object, pred.var = "Petal.Length", which.class = 3, 
                 prob = TRUE, train = trn, ...)
  
  # Bind together into a single data frame
  pd <- rbind(cbind(pd1, "Class" = "1"),
              cbind(pd2, "Class" = "2"),
              cbind(pd3, "Class" = "3"))
  
  # Display all three plots
  ggplot(pd, aes(x = Petal.Length, y = yhat, color = Class)) +
    geom_line() +
    xlab("Petal length") +
    ylab("Probability") +
    ggtitle(deparse(substitute(object))) +
    theme_light()
  
}
```


# Discriminant analysis

## Package: `MASS`

```{r}
iris.lda <- MASS::lda(Species ~ ., data = iris)
irisTest(iris.lda)
```

```{r}
iris.qda <- MASS::qda(Species ~ ., data = iris)
irisTest(iris.qda)
```

## Package: `mda`


# Decision trees

## Package: `rpart`

```{r}
iris.rpart <- rpart::rpart(Species ~ ., data = iris)
irisTest(iris.rpart)
```

## Package: `C50`
```{r}
set.seed(101)
iris.C5.0 <- C50::C5.0(Species ~ ., data = iris, trials = 100)
irisTest(iris.C5.0)
```

## Package: `party

```{r}
iris.ctree <- party::ctree(Species ~ ., data = iris)
irisTest(iris.ctree)
```

## Package: `partykit

```{r}
iris.ctree2 <- partykit::ctree(Species ~ ., data = iris)
irisTest(iris.ctree2)
```


# Bagging

## Package: `adabag`

```{r}
set.seed(101)
iris.bagging <- adabag::bagging(Species ~ ., data = iris)
irisTest(iris.bagging)
```

## Package: `ipred

```{r}
set.seed(101)
iris.ipred <- ipred::bagging(Species ~ ., data = iris, nbagg = 500)
irisTest(iris.ipred)
```

# Random forests

## Package: `randomForest

```{r}
set.seed(101)
iris.rf <- randomForest::randomForest(Species ~ ., data = iris)
irisTest(iris.rf)
```


## Package: `party

```{r}
set.seed(101)
iris.crf <- party::cforest(Species ~ ., data = iris)
irisTest(iris.crf, quantiles = TRUE)
```

## Package: `partykit`
```{r}
set.seed(101)
iris.crf2 <- partykit::cforest(Species ~ ., data = iris)
irisTest(iris.crf2, quantiles = TRUE)
```

## Package: `ranger`

```{r}
set.seed(101)
iris.ranger <- ranger::ranger(Species ~ ., data = iris, probability = TRUE)
irisTest(iris.ranger)
```


# Boosting

## Package: `adabag`

```{r}
set.seed(101)
iris.boosting <- adabag::boosting(Species ~ ., data = iris)
irisTest(iris.boosting)
```

## Package: `gbm`

```{r}
set.seed(101)
iris.gbm <- gbm::gbm(Species ~ ., 
                     data = iris,
                     distribution = "multinomial",
                     n.trees = 5000,
                     interaction.depth = 3,
                     shrinkage = 0.001,
                     cv.folds = 5,
                     verbose = FALSE)
best.iter <- gbm::gbm.perf(iris.gbm, method = "cv", plot.it = FALSE)
irisTest(iris.gbm, n.trees = best.iter)
```

## Package: `xgboost`
```{r}
set.seed(101)
iris.xgb <- xgboost::xgboost(
  data = data.matrix(subset(iris, select = -Species)), 
  label = unclass(iris$Species) - 1, num_class = 3, 
  objective = "multi:softprob", nrounds = 100, max_depth = 3, eta = 0.1, 
  gamma = 0, colsample_bytree = 0.8, min_child_weight = 1, subsample = 0.7,
  verbose = 0
)
irisTest(iris.xgb)
```

# Neural networks

## Package: `nnet`
```{r}
set.seed(101)
iris.nnet <- nnet::nnet(Species ~ ., data = iris, size = 10, decay = 0.1, 
                        maxit = 500, trace = FALSE)
irisTest(iris.nnet)
```


# Support vector machines

## Package: `e1071`

```{r}
iris.svm <- e1071::svm(Species ~ ., data = iris, type = "C-classification", 
                       probability = TRUE)
irisTest(iris.svm)
```

## Package: `kernlab`

```{r}
iris.ksvm <- kernlab::ksvm(Species ~ ., data = iris, type = "C-svc", 
                           prob.model = TRUE)
irisTest(iris.ksvm)
```


# Linear/generalized linear models

## Package: `nnet`

```{r}
set.seed(101)
iris.multinom <- nnet::multinom(Species ~ ., data = iris, trace = FALSE)
irisTest(iris.multinom)
```
